---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  echo = T,
  width = 90,
  comment = "#>",
  fig.path = "README-"
)
options(width = 100)
```

# entbal - An alternative implementation of entropy balancing weights

The goal of entbal is to create an easy to use implementation of the entropy balancing algorithm outlined in Hainmueller (2012) for applied researchers to use.  

Hainmueller, Jens. "Entropy balancing for causal effects: A multivariate reweighting method to produce balanced samples in observational studies." Political Analysis 20.1 (2012): 25-46.

## Installation

You can install entbal from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("bvegetabile/entbal")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(entbal)
library(survey, quietly = T, warn.conflicts = F, verbose = F)
dset <- dw99cps1
estwts <- entbal(TA ~ age + education + black + hispanic + married + nodegree + RE74 + RE75,
                      data = dset,
                      n_moments = 3,
                      verbose = 1,
                      estimand = "ATT")
summary(estwts)
dset$wts <- estwts$wts
design <- svydesign(ids=~1, weights=~wts, data = dset)
resp <- svyglm('RE78 ~ TA', design = design)
summary(resp)
```
