---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  echo = T,
  width = 90,
  comment = "#>",
  fig.path = "README-"
)
options(width = 100)
```

# entbal - An alternative implementation of entropy balancing weights

The goal of entbal is to create an easy to use implementation of the entropy balancing algorithm outlined in Hainmueller (2012) for applied researchers to use.  **Note that this is a work in progress**

Hainmueller, Jens. "Entropy balancing for causal effects: A multivariate reweighting method to produce balanced samples in observational studies." Political Analysis 20.1 (2012): 25-46.

## Installation

You can install entbal from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("bvegetabile/entbal")
```

## Example of Estimating Entropy Balancing Weights - ATT

This is a basic example that shows how to obtain weights on a simple dataset from Dehejia and Wahba (1999).  

Dehejia, Rajeev H., and Sadek Wahba. "Causal effects in nonexperimental studies: Reevaluating the evaluation of training programs." Journal of the American statistical Association 94.448 (1999): 1053-1062.

In the dataset the goal is estimating postintervention earnings after inclusion into a job training program.  The variables *age*, *education*, *black*, *hispanic*, *married*, *nodegree*, *RE74*, and *RE75*, all represent pre-exposure variables.  The variables *RE74*, *RE75*, represent "retrospective earnings" in 1974  and 1975 respectively.  The variable *RE78* represents the outcomes of earnings in 1978.  Finally, the variable *TA* represents the indicator of inclusion in the treatment program.  

The original DW99 paper contains six control datasets.  The example below uses a control group from the Current Population Survey (CPS1) and is the largest dataset examined in DW99.  

```{r example}
library(entbal)
library(survey, quietly = T, warn.conflicts = F, verbose = F)
dset <- dw99cps1
estwts <- entbal(TA ~ age + education + black + hispanic + married + nodegree + RE74 + RE75,
                 data = dset,
                 n_moments = 3,
                 verbose = 1,
                 bal_tol = 1e-3,
                 estimand = "ATT")
summary(estwts)
```


### Diagnostic Plots

We can visualize the weighted ECDF as a visual balance diagnostic.  We only need to do this for continuous variables as we can see from the summary table that the binary variables are balanced. These should be the columns of estwts$X that are continuous.  

```{r, fig.width=8, fig.height=4, fig.align='center'}
plot(estwts, which_vars = c(1,2,7,8))
```

These plots demonstrate very good balance on the empirical distribution. 

### Outcome Modeling

To estimate the effect of the job training program on postintervention earnings we can use the survey package to estimate the effect of treatment.   

```{r example2}
dset$wts <- estwts$wts
design <- svydesign(ids=~1, weights=~wts, data = dset)
resp <- svyglm(RE78 ~ TA,
               design = design)
summary(resp)
```


We see that the estimating difference in the post-intervention earnings is approximately $1392.  An additional analysis could controlling for additional variables, as follows:

```{r example3}
resp <- svyglm(RE78 ~ TA + age + education + black + hispanic + married + nodegree + RE74 + RE75,
               design = design)
summary(resp)
```

We see that the results do not change much and that our point estimate is the same.

## Simulated Demo on Estimating the ATE

Below is a demonstration of estimating the average treatment effect

```{r, example4, fig.width=8, fig.height=3, fig.align='center'}
set.seed(2019)

library(entbal)
library(survey)
n_obs <- 2500
n_sims <- 500
outro <- matrix(NA, nrow = n_sims, ncol = 4)
start_time <- Sys.time()
for(i in 1:n_sims){
  
  X1 <- rnorm(n_obs)
  X2 <- rnorm(n_obs)
  X3 <- rnorm(n_obs)
  
  ps <- plogis(1.5 + 1.2 * X1 - 1.5 * X2 + 0.5 * X3)
  
  TA <- rbinom(n_obs, 1, ps)
  
  Y1 <- X1^2 + X2^2 + X3^2 + 10 + rnorm(n_obs)
  Y0 <- -(X1 + X2 + X3) + rnorm(n_obs)
  
  Yobs <- TA * Y1 + (1 - TA) * Y0
  
  dset <- data.frame('X1' = X1,
                     'X2' = X2,
                     'X3' = X3,
                     'TA' = TA,
                     'Y' = Yobs)
  
  SATE <- mean(Y1-Y0)
  
  # Naive Estimate - Only taking group means
  Naive <- mean(dset$Y[dset$TA == 1]) - mean(dset$Y[dset$TA == 0])
  
  # Linear Regression - Controlling for X1, X2, X3, - Complete Case Analysis
  LR <- lm(Y ~ TA + X1 + X2 + X3, data = dset)
  
  # Entropy Balancing ATE
  wts <- entbal(TA ~ X1 + X2 + X3 ,
                verbose = 0, n_moments = 2, estimand = 'ATE')
  dset$wts <- wts$wts
  design <- svydesign(ids=~1, weights=~wts, data = dset)
  resp <- svyglm('Y ~ TA', design = design)
  
  # Storing outputs
  outro[i,] <- c(SATE, Naive, coef(LR)[2], coef(resp)[2])
  if(!(i%%100)) {
    end_time <- Sys.time()
    message(
      paste(i, ' Iters, Time per iter:', 
            round(difftime(end_time, start_time, units = 'secs'),3),
            sep = '')
    )
    start_time <- Sys.time()
  }
}
colnames(outro) <- c('SATE',
                     'Naive Analysis',
                     'Linear Regression',
                     'Entropy Balancing, m = 3')

par(mar=c(5,12,2,2))
plot(0, ylim = c(0,5), xlim = range(outro - mean(outro[,1])),
     pch = 19, col = rgb(0,0,0,0),
     xlab = 'Estimated ATE', ylab = '', axes = F, yaxs='i',
     main = 'Distribution of Estimates - Centered on Mean SATE')
lines(c(0,0), c(0, 10), lwd = 3, col = rgb(0,0,0.5,0.5))
boxplot((outro - mean(outro[,1]))[,4:1], horizontal=T, axes = F, add = T,
        col = 'gray', pch = 19)
axis(1); axis(2, las=2, at = 4:1, labels = colnames(outro))
abline(h = c(3.5,8.5), lwd = 2, col = rgb(0.5,0,0,0.4))
```
